### 2021-04-15 draft/supp.tex
aesthetic changes (for consistency and clarity of math notations)

\documentclass[a4paper,10pt]{article}

% packages
\usepackage[utf8]{inputenc} % allows usage of spanish special characters
\usepackage[spanish,english]{babel} % english dictionary for proper hyphenation
\usepackage{amsmath} % math expressions
\usepackage{upgreek} % upright greek letters for math
\usepackage{txfonts} % nice fonts 
\usepackage{authblk} % author affiliations
\usepackage{graphicx} % images
\usepackage{float} % image positioning
\usepackage{multirow} % allows merging cells in tables
\usepackage{mathtools} % nice matrices
\usepackage{rotating} % rotated text
\usepackage{caption} % customization of figure/table captions
\usepackage{subcaption} % collages of multiple images
\usepackage{hyperref} % hyperlinks
\usepackage{nameref} % cross-references between draft and supplementary material
\usepackage{zref-xr,zref-user} % more cross-references formatting
\usepackage[square,numbers,sort&compress]{natbib}
\usepackage{xcolor} % text color
\usepackage[margin=3cm]{geometry} % margins

\usepackage{pdflscape} %% Landscape
\usepackage{dcolumn}
\usepackage{booktabs}
\usepackage{savesym}
\savesymbol{iint}
\restoresymbol{TXF}{iint}
%% \usepackage{pdfpages} %% to include a pdf file

% customization
\setcounter{Maxaffil}{0} % affiliations
\renewcommand\Affilfont{\itshape\small} % style of affiliations text
\makeatletter\renewcommand{\@biblabel}[1]{#1.}\makeatother % change [number] for number. in reference list
\newcommand{\idea}[1]{\textcolor{red}{#1}}
\newcommand{\mr}{\textcolor{red}{\textbf{[missing ref(s)]}}}
\newcommand{\Burl}[1]{\url{#1}}
\zxrsetup{toltxlabel} % cross references: draft & suppl mat
\zexternaldocument*[M-]{draft}
\hypersetup{
  colorlinks = true,
  citecolor=  black,
  linkcolor = {blue},
  filecolor = cyan % controls color of external ref, if used
}
\renewcommand{\thefigure}{\textbf{S\arabic{figure}}} % preceed the Suppl Mat figures with S
\renewcommand{\thetable}{\textbf{S\arabic{table}}} % preceed the Suppl Mat tables with S
\renewcommand{\theequation}{S\arabic{equation}} % preceed the Suppl Mat equations with S
\captionsetup[figure]{font=small, labelfont={bf}, labelformat={default},
   labelsep=period, name={Fig.}} % custom figure captions
\captionsetup[table]{font=small, labelfont={bf}, labelformat={default},
   labelsep=period, name={Table}} % custom table captions

% title, authors, affiliations
\title{Supplementary Material for ``Provisional Title'' }
\author[1,2]{Authors}
\author[1,$\dagger$]{Álvaro Sánchez}
\affil[1]{Department of Ecology \& Evolutionary Biology and Microbial
  Sciences Institute, Yale University, New Haven, CT, USA}
\affil[2]{Other affiliations...}
\affil[$\dagger$]{To whom correspondence should be addressed: \normalfont alvaro.sanchez@yale.edu}
\date{}

\newcommand\legendtreefigs{Internal nodes show the partitioning variables   and the values that lead to each partition; light blue denotes the   values or categories that lead down the left split and salmon the values   or categories that lead down the right split (partitioning criteria are   also shown in the edges leading down a node). Leaves or terminal nodes   show (weighted) boxplots of the dependent variable for the observations   that belong that fall in that leave and, as label, the name of the node   ax_3,nd the sample size (sum of weighted observations and \% of the total   data set for the fit). The dot plots underneath the boxplots show the   mean JS for all the methods for that leave.}

\begin{document}

\maketitle
\tableofcontents
\newpage

\section{Supplementary Methods}\label{supp-methods}

\subsection{Data processing}\label{supp-methods:data-proc}

Thresholds for vector similarity in ESV space, etc.

\subsubsection{Special cases}\label{supp-methods:data-proc:special}

Blablabla.

\subsection{Inferring species abundance from sequencing data}
\label{supp-methods:species-from-seq}

Consider a community of $N$ species with relative abundances represented
by the components of a vector $\mathbf{x} = \left( x_1, x_2, \dotsc,
x_N\right)$ such that $x_i$ is the relative abundance of the $i$-th species.
Naturally, the conditions
%
\begin{equation}
\sum_i x_i = 1
\label{supp-eqn:normx_1}
\end{equation}
%
and
%
\begin{equation}
0 \leq x_i \leq 1 \; \textrm{for all} \; i
\label{supp-eqn:normx_2}
\end{equation}
%
are satisfied.

Performing 16S rRNA gene sequencing on such a community yields a list of
sequences (exact sequence variants or ESVs) and their respective abundance.
Normalizing by the total number of sequences, we can obtain a vector $\mathbf{s}
= \left( s_1, s_2, \dotsc \right)$ that also satisfies the normalization
conditions of equations \ref{supp-eqn:normx_1} and \ref{supp-eqn:normx_2}.
Only if there are as many ESVs as species does $\mathbf{s}$ have length $N$
(the same as $\mathbf{x}$), but in the most general case it is possible that
\textit{a)} multiple species share a same ESV and/or \textit{b)} species carrying
multiple copies
of the 16S rRNA gene have different sequences for each copy\mr. We denote
the length of $\mathbf{s}$ as $M$, which need not be equal to $N$.

By sequencing each species individually, we can build a $M \times N$ 
matrix $\mathbf{Q}$ such
that the element in the $(i,j)$ position, $q_{ij}$, represents the frequency of
the $i$-th ESV when the $j$-th species is sequenced. $\mathbf{Q}$ can be used
to determine the fraction of sequences of a given ESV that are obtained from
sequencing a community with any arbitrary composition. In other words,
$\mathbf{Q}$ maps $\mathbf{x}$ to $\mathbf{s}$:

\begin{equation}
\mathbf{Q} \cdot \mathbf{x} = \mathbf{s}
\label{supp-eqn:Qxs}
\end{equation}

Note that equation \ref{supp-eqn:Qxs} is only true if all species have the same
number of copies of the 16S rRNA gene. Otherwise, species with high copy numbers
will yield more sequences, thus leading to an over-representation and vice-versa.
Information on the 16S copy number for every species in the communities would be
required to overcome this limitation, but this is not feasible when dealing with
highly diverse communities. For our purpose in this
work, we will need to assume that the copy numbers of the 16S are relatively
conserved across species and equation \ref{supp-eqn:Qxs} holds. Additionally,
we have sequencing data from a subset of the species in our
communities, but not from all of them. This means that we cannot build a complete
$\mathbf{Q}$ matrix. To address this, we first identify every ESV obtained
from community sequencing that we cannot map back to (at least) one of the
species for which we have single-species sequencing information. We then
assume that each
of those ESVs maps uniquely to a single species. This gives us a $\mathbf{Q}$
matrix of the form:

\begin{equation}
\mathbf{Q} = \; \; \; \; \; \begin{matrix}
 &

\begin{matrix}
\; \; \;
\rotatebox{90}{\scriptsize species 1} \; \; \; & 
\rotatebox{90}{\scriptsize species 2} \; \; \; &
\rotatebox{90}{\scriptsize species 3} \; \; &
\cdots \; \; &
\rotatebox{90}{\scriptsize species $N-2$} \! \! & 
\rotatebox{90}{\scriptsize species $N-1$} \! \! &
\rotatebox{90}{\scriptsize species $N$}
\end{matrix}

\\

\begin{matrix*}[r]
\rotatebox{0}{\scriptsize sequence 1} \\
\rotatebox{0}{\scriptsize sequence 2} \\
\rotatebox{0}{\scriptsize sequence 3} \\
\vdots \; \; \; \\
\rotatebox{0}{\scriptsize sequence $M-2$} \\ 
\rotatebox{0}{\scriptsize sequence $M-1$} \\
\rotatebox{0}{\scriptsize sequence $M$}
\end{matrix*}

&

\begin{pmatrix}
q_{11} & q_{12} & q_{13} &  & 0 & 0 & 0 \\
q_{21} & q_{22} & q_{23} & \cdots & 0 & 0 & 0 \\
q_{31} & q_{32} & q_{33} &  & 0 & 0 & 0 \\
 & \vdots &  & \ddots  &  &  &  \\
0 & 0 & 0 &  & 1 & 0 & 0 \\ 
0 & 0 & 0 &  & 0 & 1 & 0 \\ 
0 & 0 & 0 &  & 0 & 0 & 1 
\end{pmatrix}

\end{matrix}
\label{supp-eqn:Q_modified}
\end{equation}
%
While a one-to-one mapping between species and ESVs may not be true in all cases,
in practice the ESVs for which we make this assumption have low relative
abundances in the communities, minimizing the impact of any potential artifact.

Having built our $\mathbf{Q}$ matrix according to equation
\ref{supp-eqn:Q_modified}, obtaining species abundances from sequencing data
is relatively straigthforward: we need to find the $\mathbf{x}$ that satisfies
equation \ref{supp-eqn:Qxs} for a given $\mathbf{s}$. However, in some cases
sequencing error can introduce deviations in $\mathbf{s}$ that make it so the
vector $\mathbf{x}$ that solves equation \ref{supp-eqn:Qxs} does not meet the
boundary conditions in equation \ref{supp-eqn:normx_2}, i.e., some $x_i$ may be
negative or greater than 1. These cases are obviously problematic, but we can
avoid them by accounting for potential deviations in $\mathbf{s}$. We do this
by solving the following nonlinear optimization problem: we want to obtain an
estimate of the true community composition
(we will denote the estimate as $\mathbf{\hat{x}}$) so that the product
$\mathbf{Q} \cdot \mathbf{\hat{x}}$ is as close as possible to the $\mathbf{s}$
obtained from sequencing while
having $\mathbf{\hat{x}}$ satisfy the normalization condition in equation
\ref{supp-eqn:normx_1} and the boundary conditions in equation
\ref{supp-eqn:normx_2}. For every possible $\mathbf{\hat{x}}$, we can define a vector
$\boldsymbol{\upvarepsilon} = \boldsymbol{\upvarepsilon} \left( \mathbf{\hat{x}} \right) = 
\left( \varepsilon_1, \varepsilon_2, \cdots, \varepsilon_M \right)$ as

\begin{equation}
\boldsymbol{\upvarepsilon} \left( \mathbf{\hat{x}} \right) =
\mathbf{Q} \cdot \mathbf{\hat{x}} - \mathbf{s}
\end{equation}
%
and a function $f \left( \mathbf{\hat{x}} \right)$ as

\begin{equation}
f \left( \mathbf{\hat{x}} \right) =
\left| \boldsymbol{\upvarepsilon} \left( \mathbf{\hat{x}} \right) \right| ^2 =
\sum_i \varepsilon_i^2 \left( \mathbf{\hat{x}} \right)
\end{equation}
%
We can also define a function $\mathbf{h} \left( \mathbf{\hat{x}} \right)$ as

\begin{equation}
\mathbf{h} \left( \mathbf{\hat{x}} \right) = 1 - \mathbf{\hat{x}}
\end{equation}
%
The best estimation for the species composition of a community will be the
$\mathbf{\hat{x}}$ that minimizes $f \left( \mathbf{\hat{x}} \right)$ while
satisfying the normalization condition (eq. \ref{supp-eqn:normx_1}) and the
boundary conditions (eq. \ref{supp-eqn:normx_2}). We solve
this problem using the \texttt{nloptr} package for R\mr, passing it
$f \left( \mathbf{\hat{x}} \right)$ as the function to minimize,
$\mathbf{h} \left( \mathbf{\hat{x}} \right) = 0$ as an equality constraint,
and 0 and 1
as the lower and upper bounds for the entries of $\mathbf{\hat{x}}$. Further
details can be found in the code for the analysis (see section 
\nameref{supp-datacode} of this Supplementary Material).

\clearpage

\section{Data \& code availability}
\label{supp-datacode}

Data and code for the analyses in this article, as well as code for the
\texttt{community-simulator} package updated with the new functionalities,
is available at \url{https://github.com/jdiazc9/coalescence}.

\clearpage

\section{Supplementary Results}\label{supp-results}

Blabla

\subsection{Example subsection}
\label{supp-results:example}

Blabla

\clearpage

\section{Supplementay Figures}\label{supp-figs}

Blablabla.

\clearpage

% references
%{\small
  %% \bibliography{refs}
%  \bibliography{extracted-supp}
% \bibliographystyle{unsrt}
%% \bibliographystyle{mystyle}
%\bibliographystyle{mybibwithurl}
%}

\end{document}